---
title: "Toronto may recover from Coronavirus if the young avoid go out"
author: "Peihan Gong"
date: 2020-09-24
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---
# Abstract:
Not only Toronto but the whole world is suffering from the Coronavirus; hence we are thinking about under this challenging situation, is there anything the public can do to reduce the probability of being infected. This paper explores the dataset of COVID-19 Cases in Toronto in 2020, which contains all confirmed and probable cases and is managed by Toronto Public Health since the first case was reported in January 2020. We find that most patients above 80 years old are infected from Toronto health care settings while the young are mostly infected from close contact. Our findings are highly instructive for future epidemic related policies by implementing specific policy on a particular age group.


# Introduction: 
Coronavirus is a hard problem facing by the whole world and seriously impact our daily lives. As of August 19, 2020, there have been approximately 123,500 confirmed cases and over 9,000 deaths in the Canada. The Government predicted around 11,000–22,000 future deaths over the pandemic, assuming "stronger epidemic control" (Wikipedia contributors. (2020, September 26)). In this paper, we aim to investigate what kind of people may have a higher rate of Coronavirus infection. In other words, is the rate of infection is uncommonly high for a group of people with a specific living habit or character? Our study is focused on gender, age, source of infection, and a development in the relationship between age group and infection source. 

Our analysis is based on 16419 cases in Toronto from January, 2020 to September, 2020. we initially find that gender is equally divided among all the cases. Also, most cases are between 20 to 59 years old, and the main infection source is close contact, outbreak-associated and community. Moreover, we find as age groups getting older, the source of infection is more likely to be marked as outbreak-associated and less likely to be marked as close contact, travel and community. 

This paper first investigates a particular variable, such as gender, age, and infection source. Then we develop a relationship between the age and the source of infection. However, since all of these data are observational data, which means that these data are just gathered without any control of subjects from researchers, thus it is not rigorous enough to infer a causal relationship. Besides, the data only contain the cases in Toronto, so the result cannot be applied to other states or countries. In future work, we can use other regions' data to investigate the characteristic of the infection patients further.


# Tool
code for this report ia available at:"https://github.com/PeihanGong/research1".

And we us R (R Core Team (2020)), R package vcd (David Meyer, Achim Zeileis, and Kurt Hornik (2020)), tidyverse (Wickham et al., (2019)), plotrix (Lemon, J. (2006)), dplyr (Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020)),ggplot2 (H. Wickham), readr (Hadley Wickham, Jim Hester and Romain Francois (2018)), plyr (Hadley Wickham (2011)), visdat (Tierney N (2017)), knitr(Yihui Xie (2014)) to do the analysis.
```{r setup, include=TRUE, warning=FALSE, message=FALSE}
library(grid)
library(vcd)
library(plotrix)
library(dplyr)
library("tidyverse")
library(plyr)
library("ggplot2")
library(readr)
library(knitr)
```


# Data
The COVID19-cases data is collected by Toronto Public Health and was combined and reported by the provincial communicable disease reporting system (iPHIS) and Toronto's custom COVID-19 case management system (CORES). The data would be refreshed every week. The data we use in this paper contains the demographic, geographic, and severity information of confirmed and probable cases from the first case in January 2020 to the nearest case on September 16, 2020. And the data set is read from a local file downloaded from Toronto Open Data portal.
```{r, warning=FALSE, message=FALSE}
COVID19_cases <- read_csv("COVID19 cases.csv")
attach(COVID19_cases)

```
Varibales and some specific word we use in this repoet and their decription are shown in the table below:

(code in appendices)
```{r echo=FALSE}
destable<-data.frame(vari_tem = c("Age Group", "Source of Infection",
                                  "Travel", "Close contact ",
                                  "Institutional setting:", "Community",
                                  "oubreak associated"),
                     descrip = c("Age at time of illness", 
                                    "most likely way that sporadic cases acquired their COVID-19 infection",
                                    "Travel outside of Ontario",
                                    "Was in close contact with a confirmed or probable COVID-19 case",
                                    "care home, hospital, etc.",
                                    "Cases with no reported travel outside of Ontario, no known close contact with a COVID-19 case, and no reported infection acquired in an institutional or healthcare setting",
                                 "associated with outbreaks of COVID-19 in Toronto healthcare institutions and healthcare settings and other Toronto congregate settings "))
colnames(destable)<-c("word we use", "description")
kable(destable)
```
Below is a graph that shows the type and missing values in the COVID 19 data set. As we can see from the graph, there are only few missing values in the variable we use, which means that it would not seriously affect our analysis.
```{r}
visdat::vis_dat(COVID19_cases)
```
The variable we focus on is gender, age and the source of infection.
Summary data are displayed by the graphs which are shown as follows:

```{r}
gen<-c(sum((dplyr::count(dplyr::filter(COVID19_cases,
                                       `Client Gender` == "FEMALE")))),
       sum((dplyr::count(dplyr::filter(COVID19_cases, 
                                       `Client Gender` == "MALE")))),
       sum((dplyr::count(dplyr::filter(COVID19_cases,
                                       `Client Gender` == "OTHER")))),
       sum((dplyr::count(dplyr::filter(COVID19_cases,
                                       `Client Gender` == "TRANSGENDER")))),
       sum((dplyr::count(dplyr::filter(COVID19_cases, 
                                       `Client Gender` == "UNKNOWN")))))
gen_word<-c("female","male","Other","transgender","unknown")
pct_1 <- round((gen/sum(gen))*100, 3)
Gender <- paste(gen_word, " ", pct_1, "%", sep = " ")

data_pie<-data.frame(gen_word,gen)

ggplot(data_pie, aes(x="", y=gen, fill=Gender)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0)+
  ggtitle("fig1:distribution of gender among COVID-19 
                infection patients in Toronto ")+
  labs(caption = "Source: Toronto Open Data Portal")+
  theme_void()


```
fig1:The graph above shows the distribution of gender among COVID -19 infection patients in Toronto. The proportion of males and females is almost the same. The proportion of female infected Coronavirus is about eight percentage points higher than male.
```{r}
COVID19_cases %>% 
  ggplot(aes(x = `Age Group`)) +
  geom_bar(fill = "skyblue",alpha = 0.8)+
  labs(x = "age group",
       y = "number of people" ,
       title = "fig2:distribution of age among COVID-19 infection patients in Toronto",
       caption = "Source: Toronto Open Data Portal" ) +
  theme(axis.text.x=element_text(angle=30, hjust=1))
```
fig2:The graph above shows the distribution of age group among COVID -19 infection patients in Toronto. Most of the patients are between 20 to 59 years old. Teenagers and the elderly are only a small proportion of the patients.
```{r}
COVID19_cases %>% 
  ggplot(aes(x = `Source of Infection`)) +
  geom_bar(fill = "skyblue", alpha = 0.8)+
  labs(x = "Source of Infection",
       y = "number of people" ,
       title = "fig3:distribution of source of infection among COVID-19 infection 
                            patients in Toronto ",
       caption = "Source: Toronto Open Data Portal") +
  theme(axis.text.x=element_text(angle=30, hjust=1))
```
fig3:The graph above shows the distribution of source of infection among COVID -19 infection patients in Toronto. Two main sources of infection are close contact and outbreak-associated. The third-largest source of infection is the community, which is only half of either the close contact and outbreak-associated. In contrast, health care and travel make up only a small percentage of the cases.


Since all the gender has an equal possibility in infecting Coronavirus, so we tried to find out is people in a particular age group is easier to infect Coronavirus from a specific source, and the analysis is shown as follows:

```{r}
counts_3<-table(COVID19_cases$`Source of Infection`,COVID19_cases$`Age Group`)
barplot(counts_3,
        main = "fig4:distribution of source of infection in each 
        age group in Toronto",
        xlab = "Age Group", ylab = "people",
        col = c("slategray2","gray","cornflowerblue","black","white",
                "yellow","blue","orange"),
        legend = rownames(counts_3))
```
fig4:The graph above shows the distribution of the source of infection in each age group in Toronto. Among all the age groups, close contact is the main cause of being infected, while outbreak-associated is another main reason for the age group above 20. For the younger age group, close contact has a larger proportion than the outbreak-associated. However, as the age group getting older, the proportion of outbreak-associated increases. 


# Discussion
Coronavirus is a hard problem facing by the whole world, and seriously impact our daily life. We cannot cure patients or develop a vaccine to stop the spread of Coronavirus but we can find the common points of the people who suffer from the Coronavirus and help the public avoid being infected.

By analyzing the data we find that both male and female has a equal chance of being infected.In other words, gender would not affect the probability of infecting a Coronavirus. Besides, the main group of people suffer from Coronavirus are the adult age from 20 to 59 and the main source of infection is close contact, outbreak association and community. In addition,We further analyze the source of infection in each of the age group. We find as the age group getting older, less people in that group are being infected from close contact,community and travel while more and more people are being marked as outbreak associated.

Since the people who is marked as outbreak associated may mainly being infected from health care settings (e.g. retirement homes, hospitals, etc.)and congregate setting (such as homeless shelters) and we cannot forbidden people especially the elderly go to hospital, thus we can only suggest other people avoid unnecessary contact with others and the best way to reduce contact is to stay at home.


## Weakness and Next Step
- In this research, we find an association with the source of infection and people's age. However, there may be a third variable (a confounding variable) affect the age group and/or source of infection. For example, the living habit may be a confounding variable. 
- The analysis is only based on the cases in Toronto, more data is needed in order to apply the result to other places. - 
- outbreak associated means the cases are associated with outbreaks of COVID-19 in Toronto healthcare institutions and healthcare settings and other Toronto  congregate settings . Thus the infection source of the cases are marked as out break associated may have many different situations but we do not know any detail. Thus, further research have to be done to deal with the problem.



# Reference
- David Meyer, Achim Zeileis, and Kurt Hornik (2020). vcd: Visualizing Categorical Data. R package version 1.4-8.
-Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data Manipulation. R package version 0.8.5. https://CRAN.R-project.org/package=dplyr
- Hadley Wickham (2011). The Split-Apply-Combine Strategy for Data Analysis. Journal of Statistical Software, 40(1), 1-29. URL http://www.jstatsoft.org/v40/i01/.
- H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
- Hadley Wickham, Jim Hester and Romain Francois (2018). readr: Read Rectangular Text Data. R package version 1.3.1. https://CRAN.R-project.org/package=readr
- Lemon, J. (2006) Plotrix: a package in the red light district of R. R-News, 6(4): 8-12.
- R Core Team (2020). R: A language and environment for statistical computing. R 
Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
- Tierney N (2017). “visdat: Visualising Whole Data Frames.” _JOSS_, *2*(16), 355. doi:
10.21105/joss.00355 (URL: https://doi.org/10.21105/joss.00355), <URL:
http://dx.doi.org/10.21105/joss.00355>.
- Toronto Public Health. (2020, September 9). Open Data Dataset. City of Toronto Open Data     Portal. https://open.toronto.ca/dataset/covid-19-cases-in-toronto/
- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
- Wikipedia contributors. (2020, September 26). COVID-19 pandemic in Canada. Wikipedia. https://en.wikipedia.org/wiki/COVID-19_pandemic_in_Canada
- Yihui Xie (2014) knitr: A Comprehensive Tool for Reproducible Research in R. In Victoria Stodden, Friedrich 
Leisch and Roger D. Peng, editors, Implementing Reproducible Computational Research. Chapman and Hall/CRC.
ISBN 978-1466561595

# Appendices

```{r,echo=TRUE, results='hide'}
destable<-data.frame(vari_tem = c("Age Group", "Source of Infection",
                                  "Travel", "Close contact ",
                                  "Institutional setting:", "Community",
                                  "oubreak associated"),
                     descrip = c("Age at time of illness", 
                                    "most likely way that sporadic cases acquired their COVID-19 infection",
                                    "Travel outside of Ontario",
                                    "Was in close contact with a confirmed or probable COVID-19 case",
                                    "care home, hospital, etc.",
                                    "Cases with no reported travel outside of Ontario, no known close contact with a COVID-19 case, and no reported infection acquired in an institutional or healthcare setting",
                                 "associated with outbreaks of COVID-19 in Toronto healthcare institutions and healthcare settings and other Toronto congregate settings "))
colnames(destable)<-c("word we use", "description")
kable(destable)
```



